Analytic Occultation Light Curves in Reflected Light

### 2020-04-04 tex/ms.tex
slow progress

@@

        b I_\frac{\nu}{2}(b) K_{\frac{\mu}{2}, \frac{\nu}{2}}

@@

        I_{\frac{\nu + 1}{2}}(b) K_{\frac{\mu - 1}{2}, \frac{\nu + 1}{2}}

@@

\section{The Solution: Occultation}
\label{sec:solution-occ}

\begin{figure}[t!]
    \begin{centering}
        \includegraphics[width=\linewidth]{figures/cases.pdf}
        \oscaption{cases}{%
            The eight families of cases of occultations in reflected light.
            In these figures, the body with the solid outline
            is the one whose flux we are interested in, and the body with the
            dashed outline is the occultor.
            The night side of the occulted body is colored
            black (dark grey if occulted), and the dayside is colored blue
            (light grey if occulted).
            The cases in the top row involve
            configurations in which the limb of the occultor does not intersect
            with the terminator at any point, so the visible flux may be
            computed in terms of classical \starry integrals. The remaining
            cases require integration along the red boundary (the curves
            $\mathcal{P}$, $\mathcal{T}$, and $\mathcal{Q}$ of
            \S\ref{sec:cases-5-8}), which
            include the terminator. These involve the evaluation of incomplete
            elliptic integrals and are derived below.
            \label{fig:cases}
        }
    \end{centering}
\end{figure}

The integration in the unocculted case presented above is relatively
straightforward, since
the boundaries of integration are always the half-ellipse defining the
terminator and the half-circle defining the upper limb of the body
(Equation~\ref{eq:rT}). When an occultor is present, however, the
integration boundaries are far less trivial, since they may or may not
include sections of the terminator, sections of the limb of the body,
and sections of the limb of the occultor. There are in total eight families of
configurations, each defined by a distinct combination of integration
boundaries; these are shown in Figure~\ref{fig:cases}.
Together, these cases encompass all possible
occultation configurations, for any illumination angle, occultor size, and
occultor position.

Before we discuss how to compute the occultation integrals, we must first
develop a procedure to identify the relevant case given the occultor
impact parameter $b_o$ and radius $r_o$ and the terminator semi-minor
axis $b$ and angle $\theta'$ in the frame $\mathcal{F}'$.
Then, once the case is determined, we must
identify the relevant integration boundaries, which depend on the points
of intesection between the limb of the body, the limb of the occultor, and
the terminator. We do so in the following section.

\subsection{Case determination and integration boundaries}
\label{sec:which-case}
%
The key to identifying the case corresponding to a given configuration is
to determine whether or not the limb of the occultor intersects the terminator
of the body, and if so, the points of intersection.
While we perform the integration in frame $\mathcal{F}'$, finding the points
of intersection with the terminator is easier if we temporarily switch to
the frame $\mathcal{F}''$ (via Equation~\ref{eq:R''}) in which the terminator
is parallel to the $x''$-axis.
In this frame, the equations defining the terminator and the limb of the
occultor are, respectively,
%
\begin{proof}{}
    y''_1(x'') & = b \sqrt{1 - x''^2}
    \nonumber                                               \\
    y''_2(x'') & = y''_o \pm \sqrt{r_o^2 - (x'' - x''_o)^2}
\end{proof}
%
where
%
\begin{proof}{}
    x''_o = b_o\sin\theta'
    \nonumber \\
    y''_o = b_o\cos\theta'
\end{proof}
%
are the coordinates of the occultor in $\mathcal{F}''$.
%
We wish to find the set of $N$ points
$\{{x''}_n~\big|~0~\le~n~<~N\}$
for which
$y''_1(x_n'') - y''_2(x_n'') = 0$. Following \citet{Luger2017}, we may
express this condition as the quartic equation
%
\begin{proof}{}
    \label{eq:quartic}
    A {x''}^4 + B {x''}^3 + C {x''}^2 + D {x''} + E = 0
\end{proof}
%
with coefficients
%
\begin{proof}{}
    \label{eq:quartic-coeffs}
    A &= (1 - b^2)^2
    \nonumber \\
    B &= -4 x''_o (1 - b^2)
    \nonumber \\
    C &= -2 \bigg(
    b^4
    + r_o^2
    - 3 {x''_o}^2
    - {y''_o}^2
    - b^2 \big(1 + r_o^2 - {x''_o}^2 + {y''_o}^2\big)
    \bigg)
    \nonumber \\
    D &= -4 x''_o (b^2 - r_o^2 + {x''_o}^2 + {y''_o}^2)
    \nonumber \\
    E &=
    b^4
    - 2 b^2 \big(r_o^2 - {x''_o}^2 + {y''_o}^2\big)
    + \big(r_o^2 - {x''_o}^2 - {y''_o}^2\big)^2
    \quad.
\end{proof}
%
Although closed-form solutions to quartic equations exist
\citep[see, e.g.,][who solve for the area of overlap between two ellipses
    analytically]{Hughes2011}, they are prone to significant numerical
instabilities. Instead, we solve for the roots of the quartic
numerically by casting it
as an eigenvalue problem \citep[e.g.,][]{Edelman1995} and polish the
results with a few iterations of Newton's method. We find that this is
reasonably computationally efficient and
yields roots with precision within a couple orders of magnitude of machine
epsilon (see \S\ref{sec:performance}).

@@

\xxx{More here...}

\subsection{Cases 1--4}
\label{sec:cases-1-4}
%
\xxx{This section needs to be revised.}
Cases 1--4 (see Figure~\ref{fig:cases}) involve configurations in which the
occultor does not intersect with
the terminator of the occulted body, and are therefore fairly
straightforward to solve.
%
Case 1 corresponds to any complete occultation of the body, so the
solution for the flux is trivial:
%
\begin{align}
    \label{eq:sT1}
    f_1 = 0
    \quad.
\end{align}
%
Case 2 involves any occultation in which the occultor blocks \emph{only} the
night side of the body (regardless of whether or not it intersects with the
limb of the body). Since the night side intensity is zero everywhere, this case
is also trivial, as the flux is equal to the flux in the no occultation case
(Equation~\ref{eq:f0}):
%
\begin{align}
    \label{eq:sT2}
    f_2 = f_0
    \quad.
\end{align}
%
Case 3 corresponds to occultations in which the occultor blocks \emph{all} of
the night side of the body and \emph{some} of the day side. In this
configuration, the unocculted part of the disk consists only of dayside.
Since we have no intersections with the terminator to worry about,
we may compute the total flux using the \starry formalism from
\citet{Luger2019}, provided we weight the intensity map by the illumination
field:
%
\begin{proof}{}
    \label{eq:f3}
    f_3 = \sTe(b_o, r_o)
    \BF{A_2}
    \BF{I}(b, \theta, r_s)
    \BF{A_1}
    \BF{R}'(\omega)
    \BF{R}(i, \lambda, \vartheta)
    \BF{y}
    \quad.
\end{proof}
%
Finally, case 4 involves any occultation in which the occultor blocks
\emph{only} the day side of the body (regardless of whether or not it
intersects with the limb):
%
\begin{proof}{}
    \label{eq:sT4}
    f_4 = f_3 + \hat{f}_0
    \quad,
\end{proof}
%
where we define
%
\begin{proof}{}
    \label{eq:f0hat}
    \hat{f}_0 & =
    \BS{\rho}^\top(-b, r_s)
    \BF{A_1}
    \BF{R}'(-\theta - \pi)
    \BF{R}(i, \lambda, \vartheta)
    \BF{y}
    \quad.
\end{proof}

@@

\subsection{Cases 5--8}
\label{sec:cases-5-8}
%
\xxx{Cases 5--8} (see Figure~\ref{fig:cases}, Figure~\ref{fig:geometry}).

% case 5: sT0 - sT0 * I
% case 6: (sTe + sT0) * I + ~sT0
% case 7: (sTe - sT0) * I
% case 8: sT0 * I

\begin{figure}[t!]
    \begin{centering}
        \includegraphics[width=\linewidth]{figures/geometry.pdf}
        \oscaption{geometry}{%
            Geometry of an occultation in reflected light, corresponding
            to case 5 in Figure~\ref{fig:cases}. The surface integral over
            the blue region is computed from the antiderivatives of the
            surface intensity map counter-clockwise along the boundary curves
            $\mathcal{P}$, $\mathcal{T}$, and $\mathcal{Q}$.
            \label{fig:geometry}
        }
    \end{centering}
\end{figure}

>>>>

    \label{eq:illum_poly}

@@

        b H_\frac{\nu}{2}(b) K_{\frac{\mu}{2}, \frac{\nu}{2}}

@@

        H_{\frac{\nu + 1}{2}}(b) K_{\frac{\mu - 1}{2}, \frac{\nu + 1}{2}}

@@

\section{The Solution: Occultation}
\label{sec:solution-occ}
%
The integration in the unocculted case presented above is relatively
straightforward, since
the boundaries of integration are always the half-ellipse defining the
terminator and the half-circle defining the upper limb of the body
(Equation~\ref{eq:rT}). When an occultor is present, however, the
integration boundaries are far less trivial, since they may or may not
include sections of the terminator, sections of the limb of the body,
and sections of the limb of the occultor. There are in total eight families of
configurations, each defined by a distinct combination of integration
boundaries; these are shown in Figure~\ref{fig:cases}.
Together, these cases encompass all possible
occultation configurations, for any illumination angle, occultor size, and
occultor position.

Before we discuss how to compute the occultation integrals, we must first
develop a procedure to identify the relevant case given the occultor
impact parameter $b_o = \sqrt{x_o^2 + y_o^2}$ and radius $r_o$ and
the terminator semi-minor
axis $b$ and angle $\theta'$ in the frame $\mathcal{F}'$.
Then, once the case is determined, we must
identify the relevant integration boundaries, which depend on the points
of intesection between the limb of the body, the limb of the occultor, and
the terminator. We do so in the following sections.

\subsection{Case determination}
\label{sec:which-case}

\begin{figure}[t!]
    \begin{centering}
        \includegraphics[width=\linewidth]{figures/cases.pdf}
        \oscaption{cases}{%
            The eight families of cases of occultations in reflected light.
            In these figures, the body with the solid outline
            is the one whose flux we are interested in, and the body with the
            dashed outline is the occultor.
            The night side of the occulted body is colored
            black (dark grey if occulted), and the dayside is colored blue
            (light grey if occulted).
            The cases in the top row involve
            configurations in which the limb of the occultor does not intersect
            with the terminator at any point, so the visible flux may be
            computed in terms of classical \starry integrals. The remaining
            cases require integration along the orange boundary (the curves
            $P$, $T$, and $Q$ of
            \S\ref{sec:cases-5-8}), which
            include the terminator. These involve the evaluation of incomplete
            elliptic integrals and are derived below.
            \label{fig:cases}
        }
    \end{centering}
\end{figure}

The key to identifying the case corresponding to a given configuration is
to determine whether or not the limb of the occultor intersects the terminator
of the body, and if so, the points of intersection.
While we perform the integration in frame $\mathcal{F}'$, finding the points
of intersection with the terminator is easier if we temporarily switch to
the frame $\mathcal{F}''$ (via Equation~\ref{eq:R''}) in which the terminator
is parallel to the $x''$-axis.
In this frame, the equations defining the terminator and the limb of the
occultor are, respectively,
%
\begin{proof}{}
    y''_1(x'') & = b \sqrt{1 - x''^2}
    \nonumber                                               \\
    y''_2(x'') & = y''_o \pm \sqrt{r_o^2 - (x'' - x''_o)^2}
\end{proof}
%
where
%
\begin{proof}{}
    x''_o = b_o\sin\theta'
    \nonumber \\
    y''_o = b_o\cos\theta'
\end{proof}
%
are the coordinates of the occultor in $\mathcal{F}''$.
%
We wish to find the set of $N$ points
$\{x_n''~\big|~0~\le~n~<~N\}$
for which
$y''_1(x_n'') - y''_2(x_n'') = 0$. Following \citet{Luger2017}, we may
express this condition as the quartic equation
%
\begin{proof}{}
    \label{eq:quartic}
    A {x''}^4 + B {x''}^3 + C {x''}^2 + D {x''} + E = 0
\end{proof}
%
with coefficients
%
\begin{proof}{}
    \label{eq:quartic-coeffs}
    A &= (1 - b^2)^2
    \nonumber \\
    B &= -4 x''_o (1 - b^2)
    \nonumber \\
    C &= -2 \bigg(
    b^4
    + r_o^2
    - 3 {x''_o}^2
    - {y''_o}^2
    - b^2 \big(1 + r_o^2 - {x''_o}^2 + {y''_o}^2\big)
    \bigg)
    \nonumber \\
    D &= -4 x''_o (b^2 - r_o^2 + {x''_o}^2 + {y''_o}^2)
    \nonumber \\
    E &=
    b^4
    - 2 b^2 \big(r_o^2 - {x''_o}^2 + {y''_o}^2\big)
    + \big(r_o^2 - {x''_o}^2 - {y''_o}^2\big)^2
    \quad.
\end{proof}
%
Although closed-form solutions to quartic equations exist
\citep[see, e.g.,][who solve for the area of overlap between two ellipses
    analytically]{Hughes2011}, they are prone to significant numerical
instabilities. Instead, we solve for the roots of the quartic
numerically by casting it
as an eigenvalue problem \citep[e.g.,][]{Edelman1995} and polish the
results with a few iterations of Newton's method. We find that this is
reasonably computationally efficient and
yields roots with precision within a couple orders of magnitude of machine
epsilon (see \S\ref{sec:performance}).

@@

Cases with zero roots (case 1 -- case 4) are treated in \S\ref{sec:cases-1-4},
while cases with one or two roots (case 5 -- case 8) are treated in
\S\ref{sec:cases-5-8}. Cases with three or four roots
(case 9 -- case 10) are rarely encountered
in practice, but are possible for some pathological configurations; these are
treated in \S\ref{sec:cases-9-10}.

@@

\subsection{Cases 1--4}
\label{sec:cases-1-4}
%
Cases 1--4 (see Figure~\ref{fig:cases}) involve configurations in which the
occultor does not intersect with
the terminator of the occulted body, and are therefore fairly
straightforward to solve.
%
Case 1 corresponds to any complete occultation of the body, so the
solution for the flux is trivial:
%
\begin{align}
    \label{eq:f1}
    f_1 = 0
    \quad.
\end{align}
%
Case 2 involves any occultation in which the occultor blocks \emph{only} the
night side of the body (regardless of whether or not it intersects with the
limb of the body). Since the night side intensity is zero everywhere, this case
is also trivial, as the flux is equal to the flux in the no occultation case
(Equation~\ref{eq:f0}):
%
\begin{align}
    \label{eq:f2}
    f_2 = f_0
    \quad.
\end{align}
%
Case 3 corresponds to occultations in which the occultor blocks \emph{all} of
the night side of the body and \emph{some} of the day side. In this
configuration, the unocculted part of the disk consists only of dayside.
Since we have no intersections with the terminator to worry about,
we may compute the total flux using the \starry formalism from
\citet{Luger2019}, provided we weight the intensity map by the illumination
field:
%
\begin{proof}{}
    \label{eq:f3}
    f_3 = \sTe(b_o, r_o)
    \BF{A_2}
    \BF{I}(b, \theta', r_s)
    \BF{A_1}
    \BF{R}'(x_o, y_o)
    \BF{R}(i, \lambda, \vartheta)
    \BF{y}
    \quad.
\end{proof}
%
Finally, case 4 involves any occultation in which the occultor blocks
\emph{only} the day side of the body (regardless of whether or not it
intersects with the limb). We compute this by first weighting the solution
for the occultation in emitted light by the intensity profile as in
case 3. The flux $f_3$ obtained in this fashion contains the contribution
from the entire nightside, which as we discuss above is unphysically weighted
by the negative part of Equation~(\ref{eq:illum_poly}). We can compute what
this contribution is from the complement of case 0:
%
\begin{proof}{}
    \label{eq:f0hat}
    \hat{f}_0 & =
    -\BS{\rho}^\top(-b, r_s)
    \BF{A_1}
    \BF{R}''(-x_s, -y_s)
    \BF{R}(i, \lambda, \vartheta)
    \BF{y}
    \quad,
\end{proof}
%
i.e., the flux we would get if we placed a negative illumination source
at the antipode $(-x_s, -y_s, -z_s)$ of the true source.
The total flux in this case is thus
%
\begin{proof}{}
    \label{eq:f4}
    f_4 = f_3 - \hat{f}_0
    \quad.
\end{proof}

\subsection{Cases 5--8}
\label{sec:cases-5-8}
%

\begin{figure}[t!]
    \begin{centering}
        \includegraphics[width=\linewidth]{figures/geometry.pdf}
        \oscaption{geometry}{%
            Geometry of an occultation in reflected light, corresponding
            to case 5 in Figure~\ref{fig:cases}. The surface integral over
            the occulted portion of the dayside (light gray region)
            is computed from the line integrals of the antiderivatives of the
            surface intensity map along the boundary curves
            $P$, $T$, and $Q$. See text for details.
            \label{fig:geometry}
        }
    \end{centering}
\end{figure}

Cases 5--8 (see Figure~\ref{fig:cases}) correspond to configurations in which
the limb of the occultor intersects with the terminator at either one point
(case 5) or two points (cases 6, 7, and 8). Because of these intersections, we
cannot simply re-weight the emitted light solution, as the integration
boundaries are now different. In general, we may compute the flux by
integrating the components of the Green's basis $\bg$
over the region $S$ bounded by three curves, which we denote
$P$, $T$, and $Q$. These are shown in orange
in Figure~\ref{fig:cases} and presented in more detail in
Figure~\ref{fig:geometry}.
Curve $P$ is a segment of the limb of
the occultor, parametrized by the angle $\phi \in [\phi_1, \phi_2]$;
curve $T$ is a segment of the terminator,
parametrized by the angle $\xi \in [\xi_1, \xi_2]$;
and curve $Q$ is a segment of the limb of the occulted body,
parametrized by the angle $\lambda \in [\lambda_1, \lambda_2]$.
The endpoints $\phi_1, \phi_2, \xi_1, \xi_2, \lambda_1,$ and
$\lambda_2$ are functions of the solutions to the quartic from
\S\ref{sec:which-case} and will be presented below.

As in \citet{Luger2019}, the approach to computing the surface integrals
of each of the terms in $\bg$ over $S$
is to use Green's theorem to transform them into line integrals along the
curves $P$, $T$, and $Q$. Let $\mathfrak{p}^\top$,
$\mathfrak{t}^\top$, and $\mathfrak{q}^\top$ denote the vectors
corresponding to the line integrals
along the curves $P$, $T$, and $Q$, respectively, for each of the terms in
$\bg$. The integration is performed in the frame $\mathcal{F}'$,
in a counter-clockwise direction relative to the center
of the integration region $S$.
Then the integral of $\bg$ over $S$ is%
\footnote{%
    The components of the vectors $\mathfrak{p}^\top$ and
    $\mathfrak{q}^\top$ are analogous to the ``primitive integrals''
    $\mathcal{P}$ and $\mathcal{Q}$
    defined in Equations~(30)--(32) in \citet{Luger2019}, although the
    integration limits of both and the sense of integration of $\mathcal{P}$
    are different.
}
%
\begin{align}
    \label{eq:sT}
    \sT & =
    \mathfrak{p}^\top + \mathfrak{t}^\top + \mathfrak{q}^\top
    \quad,
\end{align}
%
where the dependence on $b, \theta', b_o,$ and $r_o$ is implicit.
%
We defer the solution to Equation~(\ref{eq:sT}) to \S\ref{sec:sT} below,
as it is quite lengthy. Given $\sT$, the flux $f_S$ over the integration
region $S$ is computed from Equation~(\ref{eq:sTA2IA1RRy}):
%
\begin{align}
    f_S & =
    \mathfrak{s}^\top(b, \theta', b_o, r_o)
    \BF{A_2}
    \BF{I}(b, \theta', r_s)
    \BF{A_1}
    \BF{R}'(x_o, y_o)
    \BF{R}(i, \lambda, \vartheta)
    \BF{y}
    \quad.
\end{align}
%
Note that this is not necessarily the \emph{observed} flux, which we must
compute on a case-by-case basis below.

Case 5 corresponds to configurations in which the limb of the occultor
intersects the terminator at a single point. The integration region
(see Figures~\ref{fig:cases} and \ref{fig:geometry}) is the occulted
portion of the dayside, which is bounded by all three curves $P$, $Q$, and $T$.
The total flux may
be computed by subtracting the occulted flux $f_S$ from the total dayside
flux $f_0$:
%
\begin{proof}{}
    \label{eq:f5}
    f_5 = f_0 - f_S
    \quad.
\end{proof}
%
Cases 6--8 involve two points of intersection between the occultor limb and
the terminator. Case 6 corresponds to occultors that block some of the
nightside and some of the dayside, regardless of whether the occultor
is fully within the disk of the occulted body. We begin by computing the
flux over the unocculted region, $f_3$, which includes the spurious
nightside contribution. We then remove this contribution by noting that it
is equal to the total nightside contribution, $\hat{f}_0$, minus the
occulted nightside flux, $f_S$:
%
\begin{proof}{}
    \label{eq:f6}
    f_6 = f_3 - (\hat{f}_0 - f_S)
    \quad.
\end{proof}
%
Case 7 corresponds to occultors that \xxx{what?}
%
\begin{proof}{}
    \label{eq:7}
    f_7 = f_3 - f_S
    \quad.
\end{proof}
%
Finally, case 8 corresponds to occultors that \xxx{what?}
%
\begin{proof}{}
    \label{eq:f8}
    f_8 = f_S
    \quad.
\end{proof}
%

\subsection{Cases 9--10}
\label{sec:cases-9-10}
%
Pathological cases with 3 or 4 roots here.

\subsection{Computing the integrals $\sT$}
\label{sec:sT}
\xxx{This is going to be be painful.}
